### 2017-04-26 manuscript.rmd
fleshed out ZooidTypeEvolution section

In our previous siphonophore phylogenetic analysis [@Dunn:2005dy] there were several characters for which their evolutionary history was left equivocal, due to unresolved relationships between physonects. With our current cladistic resolution, we were able to determine the following results:

@@

Haeckel XXX noticed for the first time that some siphonophores were monoeicous and others were dioiecious. Our analyses in 2005 reconstructed this character and found a great amount of phylogenetic conservatism, with an unambiguous resolution of the MRCA as dioecious, and the appearance of monoecy in several taxa and clades (including Calycophorae) within the polytomy. Figure X shows the evolution of sex distribution in siphonophores under the current better-resolved tree model, and it strongly indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids (sensu lato). There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of Calycophorae, with a subsequent derived shift back to dioecy in the *Marrus-Erenna* clade.

@@

>>>>

In our previous siphonophore phylogenetic analysis [@Dunn:2005dy] there were several characters left with equivocal evolutionary histories, due to unresolved relationships between physonects. With our current cladistic resolution, we were able to determine the following results:

@@

[Citation needed] noticed for the first time that some siphonophores were monoeicous and others were dioiecious. Our analyses in 2005 reconstructed this character and found a great amount of phylogenetic conservatism, with an unambiguous resolution of the MRCA as dioecious, and the appearance of monoecy in several taxa and clades (including Calycophorae) within the polytomy. Figure X shows the evolution of sex distribution in siphonophores under the current better-resolved tree model, and it strongly indicates that monoecy in siphonophores from a dioecious ancestor occured twice, in the branch leading to Calycophorae and in the branch leading to Agalmatids (sensu lato). There is a small probability for an alternative scenario featuring a single gain of monoecy before the split of Calycophorae, with a subsequent derived shift back to dioecy in the *Marrus-Erenna* clade.

@@

One of the most striking aspects of siphonophore biology is their diversity of unique zooid types. Other colonial cnidarians (such as Hydractinia) and some bryozoans (example) have been found to have up to X different zooid types [Citation1 , Citation 2]. The siphonophore genus Forskalia has 6 basic zooid types (pneumatophore, nectophore, gastrozooid, palpon, bract, and gonophore), and a total of 10 counting subtypes (4 types of bract, male & female gonophores). Diphyomorphs have more than 1 type of nectophore, while Cystonects have none. Here we reconstruct the evolutionary origins ofthe different zooid types and subtypes on the present transcriptome tree.

Nectophores are retained modified medusae Codonophora use for swimming. The nectosome is the region of the colony that develops from the nectosomal growth zone. Unlike the siphosomal growth zone, the nectosome does not bud gastrozooids, but nectophores (and in the case of *Apolemia*, also palpons). If fact, with the exception of  *Physalia physalis* (which grows small nectophores near the gonodendra), siphonophore nectophores are exclusively found on the nectosome. It is likely that the MRCA of siphonophores had a nectosome, which has lost on the branch leading to Cystonects. We cannot exclude with certainty the alternative hypothesis of a nectosome-less ancestor followed by a gain of the nectosome in the branch leading to the Codonophora. The nectosome probably arose as a duplication of the siphosome, followed by functional specialization in propelling the colony. Figure X shows the origin of the dorsal nectosome in the Agalmatidae (sensu stricto) from an ancestral ventral nectosome.

All Codonophora have a nectosome, but the number and subtypes of nectophores present varies greatly between species. As shown in Figure X, most Codonophora presents the ancestral nectosome with multiple nectophores of the same subtype. However, Calycophorans evolved a different system with just 2 nectophores of one type. This shift may be associated with the loss of the pneumatophore. Not all Calycophorans remained with this arrangement. The Hippopodidae returned to bearing multiple identical nectophores, many of which are inactive and serve functions of defense (like a shell to retract in) and buoyancy. As in the rest of Calycophorans, the Hippopodids only use 2 nectophores to propel the colony. Another interesting shift occurs in the branch leading to Diphyomorpha, where the 2 nectophores specialize into 2 subtypes, associated with a shift into a vertically aligned position and pointed bell shapes. The 2 types function together in a coupled hydrodynamic system that allows very fast escape responses (Mackie 1964).

Bracts also present a complex evolutionary history of subtypes (Figure X). Bracts are highly reduced zooids unique to siphonophores, but they are only present in the Codonophora. As with the nectosome, we have ambiguity determining whether the MRCA of siphonophores had bracts or not. The MRCA of Codonophora ahd only one bract subtype, which was lost in Hippopodidae and independently evolved into 2 subtypes in *Erenna* and *Nanomia*. During the branch leading to Forskalidae, bract subtypes evolved from 1 to 4 morphologically distinct forms. Bracts are important for protection of the delicate zooids and to help maintain neutral buoyancy. Some calycophorans are able to actively regulate the Na/NH4 balance to adjust their buoyancy (ref).

The ancestral siphonophore certainly had a pneumatophore (Figure X). This zooid type helps the colony float, maintain its pose, and regulate geotaxis (Church, 2013). Despite its diverse biological functions, it is lost in Calycophorae and never gained again in that clade. Calycophorans rely on the ionic balance of their gelatinous nectophores and bracts to retain posture and neutral buoyancy.

Palpons are modified mouthless gastrozooids used for digestion and circulation of the gastrovascular fluid. They were present in the MRCA of siphonophores (Figure X), retained in most species, but lost twice independently in the branches leading to *Bargmannia* and Calycophorae. These taxa must have found other avenues to effectively circulate nutrients across the colony.
